---
title: "Leaching experiments"
author: "Dave Costello"
date: "6/4/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(boot)
library(MASS)
library(lme4)
library(AER)

setwd("/Users/dcostel3/Desktop/Bioretention/2018 analysis")
```

###References for Tobit/censored regression
Tobit/censored regression
http://seananderson.ca/2014/05/18/gamma-hurdle.html
http://www.ats.ucla.edu/stat/r/dae/tobit.htm
https://cran.r-project.org/web/packages/censReg/vignettes/censReg.pdf
http://stats.stackexchange.com/questions/149091/censored-regression-in-r (p of <DL)


## Import data
```{r}
#Non-salty tests
leach <- read.csv(file="leach.csv")

#Outliers
leach$Cu_conc[leach$Cu_conc>200] <- NA

#Removal rates
leach$Curemove <- with(leach,(Cu_in_meas-Cu_conc)/Cu_in_meas) #Calculate % removal 
leach$Curemove[leach$Curemove<0] <-0 #Set all negative "removals" (i.e., soils added Cu) to zero
leach$Znremove <- with(leach,(Zn_in_meas-Zn_conc)/Zn_in_meas) #Calculate % removal 
leach$Znremove[leach$Znremove<0] <-0 #Set all negative "removals" (i.e., soils added Zn) to zero
leach$Pbremove <- with(leach,(Pb_in_meas-Pb_conc)/Pb_in_meas) #Calculate % removal 

#Cell characteristics (from survey analysis)
soilchem <- read.csv(file="soil_metal_NEW.csv")

#Outliers
soilchem[soilchem$Sample=="30a",'Zn'] <- NA
soilchem[soilchem$Sample=="12a",'Cu'] <- NA

cell <- read.csv(file="cell_data_NEW.csv")
cell$Location_name <- as.character(cell$Location_name)

#Calculate log variables
soilchem$logOM <- log10(soilchem$OM)
soilchem$logBD <- log10(soilchem$BDnew)
soilchem$logd50 <- log10(soilchem$d50)
soilchem$logCu <- log10(soilchem$Cu)
soilchem$logZn <- log10(soilchem$Zn)
soilchem$logPb <- log10(soilchem$Pb)
cell$logDrain <- log10(cell$Drain_area)
#Convert the cell with no drain area to the mean of all others
cell$logDrain[which(is.infinite(cell$logDrain))] <- mean(cell$logDrain[is.finite(cell$logDrain)])
cell$logCell <- log10(cell$Cell_area)
cell$logDtoC <- cell$logDrain-cell$logCell

#chemmean <- aggregate(soilchem[,c('OM','BD','d50','logOM','logBD','logd50','Cu','Zn','Pb','logCu','logZn','logPb')],by=list(soilchem$Cell_no),median,na.rm=T)
#Old code that used entire soil depth, now with just surface soil chemistry
chemmean <- aggregate(soilchem[soilchem$Depth=="0 to 10",c('OM','BDnew','d50','logOM','logBD','logd50',
                                  'Cu','Zn','Pb','logCu','logZn','logPb')],
                      by=list(soilchem[soilchem$Dept=="0 to 10",]$Cell_no),median,na.rm=T)
cellmeans <- merge(cell,chemmean,by.x="Cell_no",by.y="Group.1")

newleach <- merge(leach,cellmeans[,c(1,8:10,13:28)],by="Cell_no",all.x=T,all.y=F)

#Subsets of the different metal treatments
leachH <- subset(newleach,newleach$Metal=="High")
leachL <- subset(newleach,newleach$Metal=="Low")
leachN <- subset(newleach,newleach$Metal=="No")

with(leachN, tapply(Curemove,list(Cell_no,Location),length))

#Salty water
saltleach <- read.csv(file="saltleach.csv")
saltleach$Cell_no <- factor(saltleach$Cell_no) 
summary(saltleach)
```

##Leaching statistics (mixed effects model with logit transformed removal)
###Cu leaching
```{r High metal stormwater}
plot(leachH$Curemove,ylim=c(0,1)) #No values of zero removal

plot(Curemove~Cell_no,data=leachH)
plot(logit(Curemove)~Retrofit,data=leachH)

CuH0r <- lmer(logit(Curemove)~1+(1|Cell_no),data=leachH,REML=F)
CuH0 <- lm(logit(Curemove)~1,data=leachH)
anova(CuH0r,CuH0)
  #Random effect is significant (p<0.001)

#Step 1, test for first fixed variable
candvar <- c('Age','Kfs','Retrofit','logDrain','logCell','logDtoC','logOM','BDnew','logd50','Cu','Location','logCu')

CuHresult <- data.frame("Variable"=candvar,"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- cbind(leachH[,c('Cell_no','Curemove')],leachH[,candvar[i]])
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(logit(Curemove)~(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(logit(Curemove)~Cudata[[3]]+(1|Cell_no),data=Cudata,REML=F)
  
  CuHresult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

CuHresult[order(CuHresult$Pval),] #Add Retrofit
#Age is second (p =0.017)

#Step 2, test for second fixed variable
candvar <- c('Age','Kfs','logDrain','logCell','logDtoC','logOM','BDnew','logd50',
             'Cu','logCu','Location')

CuHresult <- data.frame("Variable"=candvar,"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- cbind(leachH[,c('Cell_no','Curemove','Retrofit')],leachH[,candvar[i]])
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(logit(Curemove)~Retrofit+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(logit(Curemove)~Retrofit+Cudata[[4]]+(1|Cell_no),data=Cudata,REML=F)
  
  CuHresult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

CuHresult[order(CuHresult$Pval),] #Add BD

#Second step model
CuH1 <- lmer(logit(Curemove)~Retrofit+BDnew+(1|Cell_no),data=leachH,REML=F)
CuH2 <- lmer(logit(Curemove)~Retrofit*BDnew+(1|Cell_no),data=leachH,REML=F)
anova(CuH1,CuH2)
#Interaction is not signficant

#Step 3, test for third fixed variable
candvar <- c('Age','Kfs','logDrain','logCell','logDtoC','logBD','logd50',
             'Cu','logCu','Location')

CuHresult <- data.frame("Variable"=candvar,"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- cbind(leachH[,c('Cell_no','Curemove','Retrofit','BDnew')],leachH[,candvar[i]])
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(logit(Curemove)~Retrofit+BDnew+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(logit(Curemove)~Retrofit+BDnew+Cudata[[5]]+(1|Cell_no),data=Cudata,REML=F)
  
  CuHresult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

CuHresult[order(CuHresult$Pval),] #Add Cu

#Third step model
CuH3 <- lmer(logit(Curemove)~Retrofit+BDnew+Cu+(1|Cell_no),data=leachH,REML=F)
CuH4 <- lmer(logit(Curemove)~Retrofit*BDnew+Cu+Cu:Retrofit+(1|Cell_no),data=leachH,REML=F)
anova(CuH3,CuH4)

CuH5 <- lmer(logit(Curemove)~Retrofit*BDnew+Cu+Cu:BDnew+(1|Cell_no),data=leachH,REML=F)
anova(CuH3,CuH5)
  #None of the interactions are significant

#Step 4, test for fourth fixed variable
candvar <- c('Age','Kfs','logDrain','logCell','logDtoC','logBD','logd50',
             'logCu','Location')

CuHresult <- data.frame("Variable"=candvar,"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- cbind(leachH[,c('Cell_no','Curemove','Retrofit','BDnew','Cu')],leachH[,candvar[i]])
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(logit(Curemove)~Retrofit+BDnew+Cu+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(logit(Curemove)~Retrofit+BDnew+Cu+Cudata[[6]]+(1|Cell_no),data=Cudata,REML=F)
  
  CuHresult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

CuHresult[order(CuHresult$Pval),] #Add Cu
#Nothing else is significant

#FINAL MODEL
CuH <- lmer(logit(Curemove)~Retrofit+BDnew+Cu+(1|Cell_no),data=leachH,REML=F)
summary(CuH)
plot(CuH)

medBD <- quantile(leachH$BDnew,na.rm=T,0.5)
medCu <- quantile(leachH$Cu,na.rm=T,0.5)
lowBD <- quantile(leachH$BDnew,na.rm=T,0.1)
lowCu <- quantile(leachH$Cu,na.rm=T,0.1)

inv.logit(predict(CuH,newdata=list(Retrofit="yes",BDnew=medBD,Cu=medCu),re.form=~0))
inv.logit(predict(CuH,newdata=list(Retrofit="no",BDnew=medBD,Cu=medCu),re.form=~0))
  #Retrofits sorb less Cu

inv.logit(predict(CuH,newdata=list(Retrofit="yes",BDnew=medBD,Cu=medCu),re.form=~0))
inv.logit(predict(CuH,newdata=list(Retrofit="yes",BDnew=medBD,Cu=lowCu),re.form=~0))
  #Low Cu means less Cu sorbed

inv.logit(predict(CuH,newdata=list(Retrofit="yes",BDnew=medBD,Cu=medCu),re.form=~0))
inv.logit(predict(CuH,newdata=list(Retrofit="yes",BDnew=lowBD,Cu=medCu),re.form=~0))
  #Low bulk density means less Cu sorbed

#Find how many samples were at DL in retrofit cells
with(leachH,tapply(Curemove,Retrofit,function(X){sum(X==max(X,na.rm=T),na.rm=T)}))
with(leachH,tapply(Curemove,Retrofit,function(X){sum(!is.na(X))}))
```

RESTART HERE - THE MODEL CHANGED FOR CU
A large proportion of Cu was removed from high metal stormwater, but the removal efficiency differed between retrofit and new build biorentention cell soils (p<0.001). Retrofit cells only removed 77% of added Cu but new builds removed 90%. New builds had 22 samples <DL (out of 61) but retrofits had only 1 (out of 39). Soils with greater OM have less metal removal. In new builds, low OM (10%) removed 92% of Cu but high OM (90%) removed 84%. 

```{r Low metal stormwater}
plot(leachL$Curemove,ylim=c(0,1)) #Four values of zero
sort(leachL$Curemove)
leachL$Curemovenew <- leachL$Curemove+0.05 #Add the minimum non-zero value

plot(Curemovenew~Cell_no,data=leachL)
plot(logit(Curemovenew)~Retrofit,data=leachL)

CuL0r <- lmer(logit(Curemovenew)~1+(1|Cell_no),data=leachL,REML=F)
CuL0 <- lm(logit(Curemovenew)~1,data=leachL)
anova(CuL0r,CuL0)
  #Random effect is significant (p<0.001)

#Step 1, test for first fixed variable
candvar <- c('Age','Kfs','Retrofit','logDrain','logCell','logDtoC','logOM','logBD',
             'logd50','Cu','logCu','Location')
loc <- which(names(leachL) %in% candvar)
names(leachL)[loc]

CuLresult <- data.frame("Variable"=names(leachL)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- cbind(leachL[,c('Cell_no','Curemovenew')],leachL[,loc[i]])
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(logit(Curemovenew)~(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(logit(Curemovenew)~Cudata[[3]]+(1|Cell_no),data=Cudata,REML=F)
  
  CuLresult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

CuLresult[order(CuLresult$Pval),] #Add Retrofit

#Step 2, test for second fixed variable
candvar <- c('Age','Kfs','logDrain','logCell','logDtoC','logOM','logBD','logd50',
             'Cu','logCu','Location')
loc <- which(names(leachL) %in% candvar)
names(leachL)[loc]

CuLresult <- data.frame("Variable"=names(leachL)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- cbind(leachL[,c('Cell_no','Curemovenew','Retrofit')],leachL[,loc[i]])
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(logit(Curemovenew)~Retrofit+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(logit(Curemovenew)~Retrofit+Cudata[[4]]+(1|Cell_no),data=Cudata,REML=F)
  
  CuLresult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

CuLresult[order(CuLresult$Pval),] #Don't add anything else

#FINAL MODEL
CuL <- lmer(logit(Curemovenew)~Retrofit+(1|Cell_no),data=leachL,REML=F)
summary(CuL)
plot(CuL)

inv.logit(predict(CuL,newdata=list(Retrofit="yes"),re.form=~0))-0.15
inv.logit(predict(CuL,newdata=list(Retrofit="no"),re.form=~0))-0.15

plot(Curemove~Retrofit,data=leachL)

#Find how many samples were at DL in retrofit cells
with(leachL,tapply(Curemove,Retrofit,function(X){sum(X==max(X,na.rm=T),na.rm=T)}))
with(leachL,tapply(Curemove,Retrofit,function(X){sum(!is.na(X))}))
```

A intermediate proportion of Cu was removed from low metal stormwater, but the removal efficiency differed between retrofit and new build biorentention cell soils (p=0.018). Retrofit cells only removed 41% of added Cu but new builds removed 59%. Same pattern as what was observed for high metal stormwater.

###Zn leaching
```{r High metal stormwater}
plot(leachH$Znremove,ylim=c(0,1))
sort(leachH$Znremove) #16 values of zero removal

#leachH$Znremovenew <- leachH$Znremove+0.08
#Cannot just add the lowest value b/c cause higher values to be >1
#Do not want to delete rows because valuable information
#Change 0 to lowest values
leachH$Znremove[leachH$Znremove==0] = 0.08

plot(Znremove~Cell_no,data=leachH)

ZnH0r <- lmer(logit(Znremove)~1+(1|Cell_no),data=leachH,REML=F)
ZnH0 <- lm(logit(Znremove)~1,data=leachH)
anova(ZnH0r,ZnH0)
  #Random effect is significant (p<0.001)

#Step 1, test for first fixed variable
candvar <- c('Age','Kfs','Retrofit','logDrain','logCell','logDtoC','logOM','logBD',
             'logd50','Zn','logZn','Location')
loc <- which(names(leachH) %in% candvar)
names(leachH)[loc]

ZnHresult <- data.frame("Variable"=names(leachH)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- cbind(leachH[,c('Cell_no','Znremove')],leachH[,loc[i]])
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(logit(Znremove)~(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(logit(Znremove)~Zndata[[3]]+(1|Cell_no),data=Zndata,REML=F)
  
  ZnHresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

ZnHresult[order(ZnHresult$Pval),] #Add Retrofit

#Step 2, test for second fixed variable
candvar <- c('Age','Kfs','logDrain','logCell','logDtoC','logOM','logBD','logd50',
             'Zn','logZn','Location')
loc <- which(names(leachH) %in% candvar)
names(leachH)[loc]

ZnHresult <- data.frame("Variable"=names(leachH)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- cbind(leachH[,c('Cell_no','Znremove','Retrofit')],leachH[,loc[i]])
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(logit(Znremove)~Retrofit+(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(logit(Znremove)~Retrofit+Zndata[[4]]+(1|Cell_no),data=Zndata,REML=F)
  
  ZnHresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

ZnHresult[order(ZnHresult$Pval),] #No more variables to add


#FINAL MODEL 
ZnH <- lmer(logit(Znremove)~Retrofit+(1|Cell_no),data=leachH,REML=F)
summary(ZnH)
plot(ZnH)

inv.logit(predict(ZnH,newdata=list(Retrofit="yes"),re.form=~0))
inv.logit(predict(ZnH,newdata=list(Retrofit="no"),re.form=~0))

boxplot(Znremove~Retrofit,data=leachH)

#Find how many samples were at DL in retrofit cells
with(leachH,tapply(Znremove,Retrofit,function(X){sum(X==max(X,na.rm=T),na.rm=T)}))
with(leachH,tapply(Znremove,Retrofit,function(X){sum(!is.na(X))}))
```

Zn removal was very efficient with a lot of soils removing almost all of the 441 µg/L of added Zn. The removal rates were modified by construction type (retrofit vs. new build). Retrofits yielded less efficient metal removal (59%) than new builds (94%). Zn was below DL for 24 of 62 new build samples but only 1 of 39 retrofit samples.

```{r Low metal stormwater}
plot(leachL$Znremove,ylim=c(0,1))
sort(leachL$Znremove) #10 zeros removals

#Cannot just add the lowest value b/c cause higher values to be >1
#Do not want to delete rows because valuable information
#Change 0 to 1/2lowest values
leachL$Znremove[leachL$Znremove==0] = 0.13

plot(Znremove~Cell_no,data=leachL)

ZnL0r <- lmer(logit(Znremove)~1+(1|Cell_no),data=leachL,REML=F)
ZnL0 <- lm(logit(Znremove)~1,data=leachL)
anova(ZnL0r,ZnL0)
  #Random effect is significant (p<0.001)

#Step 1, test for first fixed variable
candvar <- c('Age','Kfs','Retrofit','logDrain','logCell','logDtoC','logOM','logBD',
             'logd50','Zn','logZn','Location')
loc <- which(names(leachL) %in% candvar)
names(leachL)[loc]

ZnLresult <- data.frame("Variable"=names(leachL)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- cbind(leachL[,c('Cell_no','Znremove')],leachL[,loc[i]])
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(logit(Znremove)~1+(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(logit(Znremove)~Zndata[[3]]+(1|Cell_no),data=Zndata,REML=F)
  
  ZnLresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

ZnLresult[order(ZnLresult$Pval),] #Nothing significant (log(Zn) is closest @ p=0.11)
sum(leachL$Znremove==max(leachL$Znremove,na.rm=T),na.rm=T)
sum(!is.na(leachL$Znremove))

#Find how many samples were at DL in retrofit cells
with(leachL,tapply(Znremove,Retrofit,function(X){sum(X==max(X,na.rm=T),na.rm=T)}))
with(leachL,tapply(Znremove,Retrofit,function(X){sum(!is.na(X))}))
```

Removal of Zn from lightly contaminated stormwater differed significantly among the bioretention cells, but it was not correlated to any of our measures of bioretention cells. 59 of 100 samples were <DL.


###Pb leaching
```{r High metal stormwater}
plot(leachH$Pbremove,ylim=c(0,1)) #No values of zero removal

plot(Pbremove~Cell_no,data=leachH)

PbH0r <- lmer(logit(Pbremove)~1+(1|Cell_no),data=leachH,REML=F)
PbH0 <- lm(logit(Pbremove)~1,data=leachH)
anova(PbH0r,PbH0)
  #Random effect is non-significant (p=0.19)

#Step 1, test for first fixed variable
candvar <- c('Age','Kfs','Retrofit','logDrain','logCell','logDtoC','logOM','logBD',
             'logd50','Pb','logPb','Location')
loc <- which(names(leachH) %in% candvar)
names(leachH)[loc]

PbHresult <- data.frame("Variable"=names(leachH)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Pbdata <- cbind(leachH[,c('Cell_no','Pbremove')],leachH[,loc[i]])
  Pbdata <- Pbdata[complete.cases(Pbdata),]
  
  MPb1 <- lmer(logit(Pbremove)~(1|Cell_no),data=Pbdata,REML=F)
  MPb2 <- lmer(logit(Pbremove)~Pbdata[[3]]+(1|Cell_no),data=Pbdata,REML=F)
  
  PbHresult$Pval[i] <- anova(MPb2,MPb1)$P[2]
}

PbHresult[order(PbHresult$Pval),] #Add Retrofit

#Step 2, test for second fixed variable
candvar <- c('Age','Kfs','logDrain','logCell','logDtoC','logOM','logBD',
             'logd50','Pb','logPb','Location')
loc <- which(names(leachH) %in% candvar)
names(leachH)[loc]

PbHresult <- data.frame("Variable"=names(leachH)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Pbdata <- cbind(leachH[,c('Cell_no','Pbremove',"Retrofit")],leachH[,loc[i]])
  Pbdata <- Pbdata[complete.cases(Pbdata),]
  
  MPb1 <- lmer(logit(Pbremove)~Retrofit+(1|Cell_no),data=Pbdata,REML=F)
  MPb2 <- lmer(logit(Pbremove)~Retrofit+Pbdata[[4]]+(1|Cell_no),data=Pbdata,REML=F)
  
  PbHresult$Pval[i] <- anova(MPb2,MPb1)$P[2]
}

PbHresult[order(PbHresult$Pval),] #Nothing else to add, logd50 is closest (p=0.14)

#FINAL MODEL 
PbH <- lmer(logit(Pbremove)~Retrofit+(1|Cell_no),data=leachH,REML=F)
summary(PbH)
plot(PbH)

boxplot(Pbremove~Retrofit,data=leachH)
boxplot(logit(Pbremove)~Retrofit,data=leachH)

inv.logit(predict(PbH,newdata=list(Retrofit="yes"),re.form=~0))
inv.logit(predict(PbH,newdata=list(Retrofit="no"),re.form=~0))

#Find how many samples were at DL in retrofit cells
with(leachH,tapply(Pbremove,Retrofit,function(X){sum(X==max(X,na.rm=T),na.rm=T)}))
with(leachH,tapply(Pbremove,Retrofit,function(X){sum(!is.na(X))}))
```

Pb removal was very efficient, with >75% of cores removing >97% of the added Pb. Retrofit cells were slightly less efficient than the new builds. Model says that new builds remove 98% of Pb but retrofits remove only 96%. 

```{r Low metal stormwater}
plot(leachL$Pbremove,ylim=c(0,1)) #No values of zero removal

#Only two values had non-complete removal. Not worth the model. 
```

Removal of Pb from low metal storwater (78 ug/L) was very complete. All but two samples removed Pb down to the DL (10 ug/L). No need to complete statistics.


###Clean stormwater additions

```{r Cu leach}
plot(leachN$Cu_conc) #Six non-zero values (out of 97 cores)

plot(Cu_conc~Cell_no,data=leachN)

CuN0r <- lmer(Cu_conc~1+(1|Cell_no),data=leachN,REML=F)
CuN0 <- lm(Cu_conc~1,data=leachN)
anova(CuN0r,CuN0)

#Cell number does not even explain any of the variation in Cu release
```

Very little Cu released when clean stormwater was added (6/97 samples). 

```{r Zn leach}
plot(leachN$Zn_conc)
sum(leachN$Zn_conc==2)/length(leachN$Zn_conc) #76% of samples had Zn < DL

plot(log(Zn_conc)~Cell_no,data=leachN)

ZnN0r <- lmer(log10(Zn_conc)~1+(1|Cell_no),data=leachN,REML=F)
ZnN0 <- lm(log10(Zn_conc)~1,data=leachN)
anova(ZnN0r,ZnN0)
#Random effect is significant (p<0.001)

#Step 1, test for first fixed variable
candvar <- c('Age','Kfs','Retrofit','logDrain','logCell','logDtoC','logOM','logBD',
             'logd50','Zn','logZn','Location')
loc <- which(names(leachN) %in% candvar)
names(leachN)[loc]

ZnNresult <- data.frame("Variable"=names(leachN)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- cbind(leachN[,c('Cell_no','Zn_conc')],leachN[,loc[i]])
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(log10(Zn_conc)~1+(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(log10(Zn_conc)~Zndata[[3]]+(1|Cell_no),data=Zndata,REML=F)
  
  ZnNresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

ZnNresult[order(ZnNresult$Pval),] #Nothing significant, logOM closest (p=0.25)


##FINAL MODEL
ZnN <- lmer(log10(Zn_conc)~1+(1|Cell_no),data=leachN,REML=F)
summary(ZnN)
plot(ZnN)
ranef(ZnN)

plot(Zn_conc~logOM,data=leachN,log="y")
```

A fair number of soils leached Zn when clean stormwater was added (24% of cores). None of our soil or site characteristics were able to predict the cells that leached Zn.


```{r Pb leach}
plot(leachN$Pb_conc) #No samples above the DL. 
```

No leach cores released Pb above the DL. No need for statistics.

###Salty stormwater comparisons
```{r salty metals Cu}
##Initial scoping models (full and minimum model)
  #Full model
CuSfull <- tobit(log(Cu_conc)~Cell_no*Location*Metal*Salt,data=saltleach,left=log(4))
  #Minimum (no variables) model
CuSnull <- tobit(log(Cu_conc)~1,data=saltleach,left=log(4))

##Forward addition of variables
##add1 checks to see if any variables should be added
##the option test="Chisq" runs a likelihood ratio test

add1(CuSnull,scope=CuSfull,test="Chisq")
CuS1 <- update(CuSnull,~.+Cell_no)

add1(CuS1,scope=CuSfull,test="Chisq")
CuS2 <- update(CuS1,~.+Metal)

add1(CuS2,scope=CuSfull,test="Chisq")
CuS3 <- update(CuS2,~.+Location)

add1(CuS3,scope=CuSfull,test="Chisq")
CuS4 <- update(CuS3,~.+Cell_no:Location)

add1(CuS4,scope=CuSfull,test="Chisq")
CuS5 <- update(CuS4,~.+Cell_no:Metal)

add1(CuS5,scope=CuSfull,test="Chisq")
#Nothing more needs to be added

CuS6 <- update(CuS5,~.+Salt) #See what happens when salt is included
add1(CuS6,scope=CuSfull,test="Chisq")

CuS7 <- update(CuS6,~.+Location:Salt)
add1(CuS7,scope=CuSfull,test="Chisq")

drop1(CuS5,test="Chisq")
#Use drop1 to double check that nothing added is now insignificant
#Here everything looks good and should stay. CuS7 is the final model

summary(CuS7)
##Plot to explain significant differences

boxplot(log(Cu_conc)~Metal*Cell_no,data=saltleach,las=2)
#Cell 21 leached in all treatments, cell 22 leached just in high, cell 23 leached rarely
boxplot(log(Cu_conc)~Location*Cell_no,data=saltleach,las=2)
#Cell 21 leached more in inflow, cell 22 leached more from drain, cell 23 rarely leached 
boxplot(log(Cu_conc)~Location*Salt,data=saltleach,las=2)
#Cell 21 leached more in inflow, cell 22 leached more from drain, cell 23 rarely leached 


saltleach$Cu_pred <- fitted(CuS5) #Predicted values from the final model
mu_Cu <- with(saltleach,tapply(Cu_pred,list(Location,Cell_no,Metal),mean))
  #Create a table of the significant variables for which we want to calculate p Cu > DL
sigma_Cu <- CuS5$scale
(p0 <- 1-pnorm((mu_Cu-log(4))/CuS5$scale))
  #This is a table of probabilities of samples below DL
```

```{r salty metals Zn}
##Initial scoping models (full and minimum model)
  #Full model
ZnSfull <- tobit(log(Zn_conc)~Cell_no*Location*Metal*Salt,data=saltleach,left=log(2))
  #Minimum (no variables) model
ZnSnull <- tobit(log(Zn_conc)~1,data=saltleach,left=log(2))

##Forward addition of variables
##add1 checks to see if any variables should be added
##the option test="Chisq" runs a likelihood ratio test

add1(ZnSnull,scope=ZnSfull,test="Chisq")
ZnS1 <- update(ZnSnull,~.+Cell_no)

add1(ZnS1,scope=ZnSfull,test="Chisq")
ZnS2 <- update(ZnS1,~.+Location)

add1(ZnS2,scope=ZnSfull,test="Chisq")
ZnS3 <- update(ZnS2,~.+Cell_no:Location)

add1(ZnS3,scope=ZnSfull,test="Chisq")
ZnS4 <- update(ZnS3,~.+Metal)

add1(ZnS4,scope=ZnSfull,test="Chisq")
ZnS5 <- update(ZnS4,~.+Cell_no:Metal)

add1(ZnS5,scope=ZnSfull,test="Chisq")
ZnS6 <- update(ZnS5,~.+Salt)

add1(ZnS6,scope=ZnSfull,test="Chisq")
#Nothing more needs to be added

drop1(ZnS6,test="Chisq")
#Use drop1 to double check that nothing added is now insignificant
#Here everything looks good and should stay. ZnS6 is the final model

summary(ZnS6)
##Plot to explain significant differences

boxplot(log(Zn_conc)~Salt,data=saltleach)
#High salt stormwater leached more metals than low salt
boxplot(log(Zn_conc)~Metal*Cell_no,data=saltleach,las=2)
#Cell 21 leached in all treatments, cell 22 leached just in high, cell 23 leached rarely
boxplot(log(Zn_conc)~Location*Cell_no,data=saltleach,las=2)
#Cell 21 & 23 leached more in inflow, cell 22 leached more from drain 

saltleach$Zn_pred <- fitted(ZnS6) #Predicted values from the final model
mu_Zn <- with(saltleach,tapply(Zn_pred,list(Location,Cell_no,Metal,Salt),mean))
  #Create a table of the significant variables for which we want to calculate p Cu > DL
sigma_Zn <- ZnS6$scale
(p0 <- 1-pnorm((mu_Zn-log(2))/ZnS6$scale))
  #This is a table of probabilities of samples below DL
```

```{r salty metals Pb}
##Initial scoping models (full and minimum model)
  #Full model
PbSfull <- tobit(log(Pb_conc)~Cell_no*Location*Metal*Salt,data=saltleach,left=log(10))
  #Minimum (no variables) model
PbSnull <- tobit(log(Pb_conc)~1,data=saltleach,left=log(10))

##Forward addition of variables
##add1 checks to see if any variables should be added
##the option test="Chisq" runs a likelihood ratio test

add1(PbSnull,scope=PbSfull,test="Chisq")
PbS1 <- update(PbSnull,~.+Cell_no)

add1(PbS1,scope=PbSfull,test="Chisq")
PbS2 <- update(PbS1,~.+Salt)

add1(PbS2,scope=PbSfull,test="Chisq")
PbS3 <- update(PbS2,~.+Metal)
#Here I added a variable even though p > 0.05 because I didn't want to miss any
#potential interactions. Using a addition threshold of p < 0.1 is fine for this
#approach to model building

add1(PbS3,scope=PbSfull,test="Chisq")
PbS4 <- update(PbS3,~.+Location)
#Same here, add p < 0.1

add1(PbS4,scope=PbSfull,test="Chisq")
PbS5 <- update(PbS4,~.+Cell_no:Location)
#This is where that became important since adding location was required to add 
#the Cell x location interaction

add1(PbS5,scope=PbSfull,test="Chisq")
#Nothing else needs to be added

drop1(PbS5,test="Chisq")
#Use drop1 to double check that nothing added is now insignificant
#Here everything looks good and should stay. PbS5 is the final model

##Plot to explain significant differences

boxplot(log(Pb_conc)~Cell_no*Location,data=saltleach)
#Cell 21 leached Pb, but only in the inflow
boxplot(log(Pb_conc)~Metal,data=saltleach)
#High Pb > Low Pb = No Pb
boxplot(log(Pb_conc)~Salt,data=saltleach)
#High salt leached more metal than low salt

saltleach$Pb_pred <- fitted(PbS5) #Predicted values from the final model
mu_Pb <- with(saltleach,tapply(Pb_pred,list(Location,Cell_no,Metal,Salt),mean))
  #Create a table of the significant variables for which we want to calculate p Cu > DL
sigma_Pb <- PbS5$scale
(p0 <- 1-pnorm((mu_Pb-log(10))/PbS5$scale))
  #This is a table of probabilities of samples below DL
```

