---
title: "Bioretention survey"
author: "Dave Costello"
date: "5/3/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Hmisc)
library(nlme)
library(lme4)
library(MASS)

setwd("/Users/dcostel3/Desktop/Bioretention/2018 analysis")
```

#Data import
```{r}
soilchem <- read.csv(file="soil_metal_NEW.csv")
summary(soilchem)

#Outliers
soilchem[soilchem$Sample=="30a",'Zn'] <- NA
soilchem[soilchem$Sample=="12a",'Cu'] <- NA
#soilchem[soilchem$Sample=="22a2",'Pb'] <- NA

cell <- read.csv(file="cell_data_NEW.csv")
cell$Location_name <- as.character(cell$Location_name)
summary(cell)

#Calculate log variables
soilchem$logOM <- log10(soilchem$OM)
soilchem$logBD <- log10(soilchem$BD)
soilchem$logd50 <- log10(soilchem$d50)
cell$logDrain <- log10(cell$Drain_area)
cell$logDrain[which(is.infinite(cell$logDrain))] <- NA
cell$logCell <- log10(cell$Cell_area)
cell$logDtoC <- cell$logDrain/cell$logCell

#Merge both datasets
fullsoil <- merge(soilchem,cell[,c(2:10,13:15)],by="Cell_no")
summary(fullsoil)

with(soilchem,tapply(OM,list(Depth,Location),length))
```

##Secondary analyses
###Correlations
```{r cell correlations}
#Calculate mean OM, BD, and d50 for cells
chemmean <- aggregate(soilchem[,c('OM','BD','d50','logOM','logBD','logd50')],
                      by=list(soilchem$Cell_no),median,na.rm=T)
cellmeans <- merge(cell,chemmean,by.x="Cell_no",by.y="Group.1")
cordata <- cellmeans[,c('Age','Cell_area','Drain_area','Kfs','LF_sus','OM','BD','d50',
                        'logOM','logBD','logd50','logDrain','logCell','logDtoC')]

#Pairwise correlations (uses Hmisc package)
pairs(cordata)
rcorr(as.matrix(cordata))

#Cell to drainage sizes
with(cordata[cordata$Cell_area<2000,],cor.test(logCell,logDrain))
plot(Cell_area~Drain_area,data=cordata[cordata$Cell_area<2000,],log="xy")

#Regular regression
rmod <- lm(logCell~logDrain,data=cordata)
summary(rmod)
#Regression testing slope = 1
rmod2 <- lm(logCell~logDrain+offset(logDrain),data=cordata)
summary(rmod2)

10^(rmod2$coef[1])

#Even without the large Mayfield Lib cite, there is a correlation between drain area and log cell area
```

Variables are not that strongly correlated to each other, which is good for multiple regression. `OM` is correlated to `BD` (neg p=0.006) and `d50` (neg p=0.016) and `BD` and `d50` are correlated (neg p=0.03). `Age` is negatively correlated to `LF_sus` (p<0.001). The relationship between `Age` and `Kfs` identified to Hartung's thesis is not significant when looking at cell averages. Log of `Drain_area` is correlated to the log of `Cell_area` (p<0.001), but the variables are not correlated on arithmetic scale. Also, cells are oversizes relative to recommendations (5-10%), with our average cell 29% of the area of impervious drainage area. Drainage and cells are proportional on a log scale (slope =1, p = 0.9) This suggests that practicioners size the biorentention cells on an order of magnitude difference relative to impervious drainage area and oversize them substantially.

###Tree lawn paired t-test
```{r tree lawn magnetics}
t.test(cell$LF_sus,cell$Adj_LF,paired=T)

t.test(cell$HF_sus,cell$Adj_HF,paired=T)
```

No difference in magnetics within biorention cells and tree lawns. HF p = 0.74, LF p = 0.71.

###Retrofit t-tests
```{r}
t.test(logBD~Retrofit,data=cellmeans)
t.test(logOM~Retrofit,data=cellmeans)
t.test(Kfs~Retrofit,data=cellmeans)
t.test(logd50~Retrofit,data=cellmeans)
t.test(LF_sus~Retrofit,data=cellmeans)
t.test(logDrain~Retrofit,data=cellmeans)
t.test(logCell~Retrofit,data=cellmeans)
t.test(logDtoC~Retrofit,data=cellmeans)
```

Bulk density is lower in new builds (p=0.004) and retrofits have slightly lower log drainage area (p=0.04). All other soil parameters do not differ.

###Inflow vs. drain and depth tests
```{r}
OMmod <- lme(logOM~Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
OMmod2 <- lm(logOM~Depth*Location,data=fullsoil)
anova(OMmod,OMmod2)
#OM was similar among all cells (random effect p=0.22)
summary(OMmod)
plot(OMmod)
#OM was lower in deeper soil (p=0.002; 10^-0.245 = 0.56) so deep soil has half the OM.
OMmod3 <- lme(logOM~Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
OMmod4 <- lme(logOM~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(OMmod3,OMmod4)
#Age doesn't improve model fit (p = 0.73)

BDmod <- lme(logBD~Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
BDmod2 <- lm(logBD~Depth*Location,data=fullsoil)
anova(BDmod,BDmod2)
#BD was similar among all cells (random effect p=0.78)
summary(BDmod)
plot(BDmod)
#BD was greater in deeper soil (p=0.008; 10^-0.29 = 0.51) so deep soil was twice as dense.
BDmod3 <- lme(logBD~Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
BDmod4 <- lme(logBD~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(BDmod3,BDmod4)
#Age doesn't improve model fit (p = 0.26)


d50mod <- lme(logd50~Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
d50mod2 <- lm(logd50~Depth*Location,data=fullsoil)
anova(d50mod,d50mod2)
#d50 differed among all cells (random effect p=0.01)
summary(d50mod)
plot(d50mod)
#d50 did not differ among depth or location
d50mod3 <- lme(logd50~1,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
d50mod4 <- lme(logd50~Age,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(d50mod3,d50mod4)
#Age doesn't improve model fit (p = 0.41)
```

Median partical size differed among the cells, but broader soil characteristics like bulk density and organic matter content were fairly similar among cells. d50, OM, and BD did not differ between the inflow and drain, but the surface soils had higher OM and lower BD (no difference in d50). Age also did not influence d50, OM, or BD.


##Zinc analyses
###Soil Zn content

```{r Zn ME for time and space}
plot(Zn~Age,data=fullsoil)

Znmod <- lme(log(Zn)~Age*Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(Znmod)
plot(Znmod)

#Simplify the model to remove NS terms to get better parameter estimates
Znmodsim <- lme(log(Zn)~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
Znmodsim2 <- lme(log(Zn)~Age+Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(Znmodsim,Znmodsim2)
summary(Znmodsim2)

#Test for significance of the random effect
Znmod_noME <- lm(log(Zn)~Age+Depth,data=fullsoil)
anova(Znmodsim,Znmod_noME)
  #Yes, retain random effect
```

Zn concentrations increase with age (p=0.0006). log(Zn) increases by 0.128*age (14% increase each year). Depth is important for Zn concentrations (p=0.027) with shallow soils having ~17% more Zn (e^0.15). Location does not matter (p=0.19).


##Other variables influencing Zn

```{r}

#Test models with new variables

#STEP 1, including one additional
candvar <- c('Cell_area','OM','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logOM','logBD','logd50','logDrain','logCell','logDtoC','Retrofit')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Znresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- fullsoil[,c(1,5:6,17,loc[i])]
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(log(Zn)~Age+Depth+(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(log(Zn)~Age+Depth+Zndata[[5]]+(1|Cell_no),data=Zndata,REML=F)
  
  Znresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

Znresult
#INCLUDE logOM (p<0.001) first.

#REFINE THE Zn MODEL
MZn2 <- lmer(log(Zn)~Age+Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
MZn2.1 <- lmer(log(Zn)~Age+Depth*logOM+(1|Cell_no),data=fullsoil,REML=F)
anova(MZn2,MZn2.1)

MZn2.2 <- lmer(log(Zn)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
anova(MZn2.2,MZn2)
  #No signficant interactions between OM and other variables

#STEP 2 - See if other variables besides OM should be included
candvar <- c('Cell_area','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logCell','logDrain','logDtoC','logBD','logd50','Retrofit')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Znresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- fullsoil[,c(1,5:6,14,17,loc[i])]
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(log(Zn)~Age+Depth+logOM+(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(log(Zn)~Age+Depth+logOM+Zndata[[6]]+(1|Cell_no),data=Zndata,REML=F)
  
  Znresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

Znresult
#No other variables are worth including.

#FINAL MODEL
MZn2.1 <- lmer(log(Zn)~Age+Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
summary(MZn2.1)
plot(MZn2.1)

MZn2.1nlme <- lme(log(Zn)~Age+Depth+log(OM),random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(MZn2.1nlme)

(newOM <- quantile(fullsoil$logOM,0.1,na.rm=T))
exp(predict(MZn2.1,newdata=list(Age=seq(0,7),Depth=rep("0 to 10",8),logOM=rep(newOM,8)),re.form=~0))

```

Zn in biorention cells is increasing with age (p<0.001), decreased with depth (p<0.001) and is greater in soils with more organic matter (`log(OM)`, p<0.001). Across the range of biorention cells that we sampled we would see approximately 34% more Zn in the high OM soils when compared to the low OM soils.


##Copper analyses
###Soil Cu content

```{r Cu ME for time and space}
plot(Cu~Age,data=fullsoil,col=as.numeric(Depth))

Cumod <- lme(log(Cu)~Age*Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(Cumod)
plot(Cumod)

#Simplify the model to remove NS terms to get better parameter estimates
Cumodsim <- lme(log(Cu)~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
summary(Cumodsim)

#Test for significance of the random effect
Cumod_noME <- lm(log(Cu)~Age*Depth,data=fullsoil)
anova(Cumodsim,Cumod_noME)
  #Yes, retain random effect
```

Cu concentrations increase with age in the surface soil but are stable in deep soils (Age * Depth, p=0.007). In surface soils, Cu concentration increase 8% year over year but are stable below 10 cm depth. 

##Other variables influencing Cu

```{r}

#Test models with new variables

#STEP 1, including one additional
candvar <- c('Cell_area','OM','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logOM','logBD','logd50','logDrain','logCell','logDtoC','Retrofit')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Curesult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- fullsoil[,c(1,3,6,17,loc[i])]
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(log(Cu)~Age*Depth+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(log(Cu)~Age*Depth+Cudata[[5]]+(1|Cell_no),data=Cudata,REML=F)
  
  Curesult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

Curesult
#INCLUDE logOM (p=0.006).

#REFINE THE Cu MODEL
MCu2 <- lmer(log(Cu)~Age*Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
MCu2.1 <- lmer(log(Cu)~Age*Depth+logOM+logOM:Depth+(1|Cell_no),data=fullsoil,REML=F)
anova(MCu2,MCu2.1)

MCu2.2 <- lmer(log(Cu)~Age*Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
anova(MCu2.2,MCu2)

MCu2.3 <- lmer(log(Cu)~Age*Depth*logOM+(1|Cell_no),data=fullsoil,REML=F)
anova(MCu2.3,MCu2)
  #No signficant interactions between OM and other variables

#STEP 2 - See if other variables besides OM should be included
candvar <- c('Cell_area','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logCell','logDrain','logDtoC','logBD','logd50','Retrofit')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Curesult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- fullsoil[,c(1,3,6,14,17,loc[i])]
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(log(Cu)~Age*Depth+logOM+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(log(Cu)~Age*Depth+logOM+Cudata[[6]]+(1|Cell_no),data=Cudata,REML=F)
  
  Curesult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

Curesult
#No other variables are worth including.

#FINAL MODEL
MCu2 <- lmer(log(Cu)~Age*Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
summary(MCu2)
plot(MCu2)

MCu2nlme <- lme(log(Cu)~Age*Depth+logOM,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(MCu2nlme)

(newOM <- quantile(fullsoil$logOM,0.5,na.rm=T))
exp(predict(MCu2,newdata=list(Age=seq(0,7),Depth=rep("0 to 10",8),logOM=rep(newOM,8)),re.form=~0))

```

Surface soils in biorentention cells increase in Cu concentrations as they age, but deeper soils have more stable Cu concentrations (p=0.009). Cu is greater in high OM soils and across the range of cells we measured (5-80% OM) there is a 50% greater concentration of Cu on high OM soils.


##Lead analyses
###Soil Pb content

```{r Pb ME for time and space}
plot(log10(Pb)~Age,data=fullsoil,col=as.numeric(Depth))

Pbmod <- lme(log(Pb)~Age*Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(Pbmod)
plot(Pbmod)

#Simplify the model to remove NS terms to get better parameter estimates
Pbmodsim <- lme(log(Pb)~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
Pbmodsim2 <- lme(log(Pb)~Age+Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(Pbmodsim,Pbmodsim2)
Pbmodsim3 <- lme(log(Pb)~Age,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(Pbmodsim2,Pbmodsim3)
summary(Pbmodsim2)
  #The interaction between depth and age is really close (p=0.07) so may want to maintain

#Test for significance of the random effect
Pbmod_noME <- lm(log(Pb)~Age*Depth,data=fullsoil)
anova(Pbmodsim,Pbmod_noME)
  #Yes, retain random effect
```

Pb concentrations increase with age, and there is marginal trend of greater increases in the surface soil when compared to deep soils (Age * Depth, p=0.07). Seems like Pb is an intermediate between Cu and Zn where Zn has no vertical stratification, Cu is strongly stratified and Pb is weakly stratified. Pb increases 4-12% (deep and surface, respectively) year-over-year.

##Other variables influencing Pb

```{r}

#Test models with new variables

#STEP 1, including one additional
candvar <- c('Cell_area','OM','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logOM','logBD','logd50','logDrain','logCell','logDtoC','Retrofit')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Pbresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Pbdata <- fullsoil[,c(1,4,6,17,loc[i])]
  Pbdata <- Pbdata[complete.cases(Pbdata),]
  
  MPb1 <- lmer(log(Pb)~Age+Depth+(1|Cell_no),data=Pbdata,REML=F)
  MPb2 <- lmer(log(Pb)~Age+Depth+Pbdata[[5]]+(1|Cell_no),data=Pbdata,REML=F)
  
  Pbresult$Pval[i] <- anova(MPb2,MPb1)$P[2]
}

Pbresult
#INCLUDE logOM (p<0.001).

#REFINE THE Pb MODEL
MPb2 <- lmer(log(Pb)~Age+Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
MPb2.1 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Depth+(1|Cell_no),data=fullsoil,REML=F)
anova(MPb2,MPb2.1)

MPb2.2 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
anova(MPb2.2,MPb2)
  #Significant age x log(OM) interaction

#STEP 2 - See if other variables besides OM should be included
candvar <- c('Cell_area','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logCell','logDrain','logDtoC','logBD','logd50','Retrofit')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Pbresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Pbdata <- fullsoil[,c(1,4,6,14,17,loc[i])]
  Pbdata <- Pbdata[complete.cases(Pbdata),]
  
  MPb1 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=Pbdata,REML=F)
  MPb2 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+Pbdata[[6]]+(1|Cell_no),data=Pbdata,REML=F)
  
  Pbresult$Pval[i] <- anova(MPb2,MPb1)$P[2]
}

Pbresult
#No other variables are worth including.

#FINAL MODEL
MPb2.2 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
summary(MPb2.2)
plot(MPb2.2)

MPb2nlme <- lme(log(Pb)~Age+Depth+logOM+Age:logOM,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(MPb2nlme)

#Predictions to try and interpret the model
newage <- seq(0,7)
(newOM <- quantile(fullsoil$logOM,0.1,na.rm=T))
highOM <- predict(MPb2.2,newdata=list(Age=newage,Depth=rep("0 to 10",8),logOM=rep(newOM,8)),re.form=~0)
lowOM <- predict(MPb2.2,newdata=list(Age=newage,Depth=rep("10 to 20",8),logOM=rep(newOM,8)),re.form=~0)

plot(Pb~Age,data=fullsoil)
lines(newage,exp(highOM))
lines(newage,exp(lowOM))
```

Pb in bioretention soils increase through time, but only in soils that had higher OM content (Age x log(OM) p = 0.0001). Low OM content soils showed no accumulation of Pb through time, but high OM soils accumuated substantial Pb (20-150 mg/kg over a 7 year period). Deep soils overall had less Pb (p=0.016) but even deep soils with high OM accumulated Pb over time.
