---
title: "Bioretention survey"
author: "Dave Costello"
date: "5/3/2018"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Hmisc)
library(nlme)
library(lme4)
library(MASS)

setwd("/Users/dcostel3/Desktop/Bioretention/2018 analysis")
```

#Data import
```{r}
soilchem <- read.csv(file="soil_metal_NEW.csv")
soilchem$Cell_no <- factor(soilchem$Cell_no)
summary(soilchem)

#Outliers
soilchem[soilchem$Sample=="30a",'Zn'] <- NA
soilchem[soilchem$Sample=="12a",'Cu'] <- NA
#soilchem[soilchem$Sample=="22a2",'Pb'] <- NA

cell <- read.csv(file="cell_data_NEW.csv")
cell$Location_name <- as.character(cell$Location_name)
summary(cell)

#Calculate log variables
soilchem$logOM <- log10(soilchem$OM)
soilchem$logBD <- log10(soilchem$BD)
soilchem$logd50 <- log10(soilchem$d50)
cell$logDrain <- log10(cell$Drain_area)
cell$logDrain[which(is.infinite(cell$logDrain))] <- NA
cell$logCell <- log10(cell$Cell_area)
cell$logDtoC <- cell$logDrain/cell$logCell

#Merge both datasets
fullsoil <- merge(soilchem,cell[,c(2:10,13:15)],by="Cell_no")
summary(fullsoil)

```

##Secondary analyses
###Correlations
```{r cell correlations}
#Calculate mean OM, BD, and d50 for cells
chemmean <- aggregate(soilchem[,c('OM','BD','d50','logOM','logBD','logd50')],
                      by=list(soilchem$Cell_no),median,na.rm=T)
cellmeans <- merge(cell,chemmean,by.x="Cell_no",by.y="Group.1")
cordata <- cellmeans[,c('Age','Cell_area','Drain_area','Kfs','LF_sus','OM','BD','d50',
                        'logOM','logBD','logd50','logDrain','logCell','logDtoC')]

#Pairwise correlations (uses Hmisc package)
pairs(cordata)
rcorr(as.matrix(cordata))

with(cordata[cordata$Cell_area<2000,],cor.test(logCell,Drain_area))
plot(Cell_area~Drain_area,data=cordata[cordata$Cell_area<2000,],log="y")
  #Even without the large Mayfield Lib cite, there is a correlation between drain area and log cell area
```

Variables are not that strongly correlated to each other, which is good for multiple regression. `OM` is correlated to `BD` (neg p=0.006) and `d50` (neg p=0.016) and `BD` and `d50` are correlated (neg p=0.03). `Age` is negatively correlated to `LF_sus` (p<0.001). Interestingly, `Drain_area` is correlated to the log of `Cell_area` (p=0.001). This suggests that practicioners don't start boosting the size of cells until the drainage area gets very large. The relationship between `Age` and `Kfs` identified to Hartung's thesis is not significant when looking at cell averages. **Check with Anne about the specs for drain:cell ratios**

###Tree lawn paired t-test
```{r tree lawn magnetics}
t.test(cell$LF_sus,cell$Adj_LF,paired=T)

t.test(cell$HF_sus,cell$Adj_HF,paired=T)
```

No difference in magnetics within biorention cells and tree lawns. HF p = 0.74, LF p = 0.71.


##Zinc analyses
###Soil Zn content

```{r Zn ME for time and space}
plot(Zn~Age,data=fullsoil)

Znmod <- lme(log(Zn)~Age*Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(Znmod)
plot(Znmod)

#Simplify the model to remove NS terms to get better parameter estimates
Znmodsim <- lme(log(Zn)~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
Znmodsim2 <- lme(log(Zn)~Age+Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(Znmodsim,Znmodsim2)
summary(Znmodsim2)

#Test for significance of the random effect
Znmod_noME <- lm(log(Zn)~Age+Depth,data=fullsoil)
anova(Znmodsim,Znmod_noME)
  #Yes, retain random effect
```

Zn concentrations increase with age (p=0.0006). log(Zn) increases by 0.12*age (13% increase each year). Depth is important for Zn concentrations (p=0.0023) with shallow soils having ~22% more Zn (e^0.20). Location does not matter (p=0.19).


##Other variables influencing Zn

```{r}

#Test models with new variables

#STEP 1, including one additional
candvar <- c('Cell_area','OM','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logOM','logBD','logd50','logDrain','logCell','logDtoC')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Znresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- fullsoil[,c(1,5:6,17,loc[i])]
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(log(Zn)~Age+Depth+(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(log(Zn)~Age+Depth+Zndata[[5]]+(1|Cell_no),data=Zndata,REML=F)
  
  Znresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

Znresult
#INCLUDE logOM (p<0.001) first.

#REFINE THE Zn MODEL
MZn2 <- lmer(log(Zn)~Age+Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
MZn2.1 <- lmer(log(Zn)~Age+Depth*logOM+(1|Cell_no),data=fullsoil,REML=F)
anova(MZn2,MZn2.1)

MZn2.2 <- lmer(log(Zn)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
anova(MZn2.2,MZn2)
  #No signficant interactions between OM and other variables

#STEP 2 - See if other variables besides OM should be included
candvar <- c('Cell_area','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logCell','logDrain','logDtoC','logBD','logd50')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Znresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Zndata <- fullsoil[,c(1,5:6,14,17,loc[i])]
  Zndata <- Zndata[complete.cases(Zndata),]
  
  MZn1 <- lmer(log(Zn)~Age+Depth+logOM+(1|Cell_no),data=Zndata,REML=F)
  MZn2 <- lmer(log(Zn)~Age+Depth+logOM+Zndata[[6]]+(1|Cell_no),data=Zndata,REML=F)
  
  Znresult$Pval[i] <- anova(MZn2,MZn1)$P[2]
}

Znresult
#No other variables are worth including.

#FINAL MODEL
MZn2.1 <- lmer(log(Zn)~Age+Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
summary(MZn2.1)
plot(MZn2.1)

MZn2.1nlme <- lme(log(Zn)~Age+Depth+log(OM),random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(MZn2.1nlme)
```

Zn in biorention cells is increasing with age (p<0.001), decreased with depth (p<0.001) and is greater in soils with more organic matter (`log(OM)`, p<0.001). Across the range of biorention cells that we sampled (5-80% OM) we would see approximately a doubling in Zn in the high OM soils when compared to the low OM soils.


##Copper analyses
###Soil Cu content

```{r Cu ME for time and space}
plot(Cu~Age,data=fullsoil,col=as.numeric(Depth))

Cumod <- lme(log(Cu)~Age*Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(Cumod)
plot(Cumod)

#Simplify the model to remove NS terms to get better parameter estimates
Cumodsim <- lme(log(Cu)~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
summary(Cumodsim)

#Test for significance of the random effect
Cumod_noME <- lm(log(Cu)~Age*Depth,data=fullsoil)
anova(Cumodsim,Cumod_noME)
  #Yes, retain random effect
```

Cu concentrations increase with age in the surface soil but are stable in deep soils (Age * Depth, p=0.007). In surface soils, Cu concentration increase 8% year over year but are stable below 10 cm depth. 

##Other variables influencing Cu

```{r}

#Test models with new variables

#STEP 1, including one additional
candvar <- c('Cell_area','OM','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logOM','logBD','logd50','logDrain','logCell','logDtoC')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Curesult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- fullsoil[,c(1,3,6,17,loc[i])]
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(log(Cu)~Age*Depth+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(log(Cu)~Age*Depth+Cudata[[5]]+(1|Cell_no),data=Cudata,REML=F)
  
  Curesult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

Curesult
#INCLUDE logOM (p=0.006).

#REFINE THE Cu MODEL
MCu2 <- lmer(log(Cu)~Age*Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
MCu2.1 <- lmer(log(Cu)~Age*Depth+logOM+logOM:Depth+(1|Cell_no),data=fullsoil,REML=F)
anova(MCu2,MCu2.1)

MCu2.2 <- lmer(log(Cu)~Age*Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
anova(MCu2.2,MCu2)

MCu2.3 <- lmer(log(Cu)~Age*Depth*logOM+(1|Cell_no),data=fullsoil,REML=F)
anova(MCu2.3,MCu2)
  #No signficant interactions between OM and other variables

#STEP 2 - See if other variables besides OM should be included
candvar <- c('Cell_area','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logCell','logDrain','logDtoC','logBD','logd50')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Curesult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Cudata <- fullsoil[,c(1,3,6,14,17,loc[i])]
  Cudata <- Cudata[complete.cases(Cudata),]
  
  MCu1 <- lmer(log(Cu)~Age*Depth+logOM+(1|Cell_no),data=Cudata,REML=F)
  MCu2 <- lmer(log(Cu)~Age*Depth+logOM+Cudata[[6]]+(1|Cell_no),data=Cudata,REML=F)
  
  Curesult$Pval[i] <- anova(MCu2,MCu1)$P[2]
}

Curesult
#No other variables are worth including.

#FINAL MODEL
MCu2 <- lmer(log(Cu)~Age*Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
summary(MCu2)
plot(MCu2)

MCu2nlme <- lme(log(Cu)~Age*Depth+logOM,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(MCu2nlme)
```

Surface soils in biorentention cells increase in Cu concentrations as they age, but deeper soils have more stable Cu concentrations (p=0.009). Cu is greater in high OM soils and across the range of cells we measured (5-80% OM) there is a 50% greater concentration of Cu on high OM soils.



##Lead analyses
###Soil Pb content

```{r Pb ME for time and space}
plot(log10(Pb)~Age,data=fullsoil,col=as.numeric(Depth))

Pbmod <- lme(log(Pb)~Age*Depth*Location,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(Pbmod)
plot(Pbmod)

#Simplify the model to remove NS terms to get better parameter estimates
Pbmodsim <- lme(log(Pb)~Age*Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
Pbmodsim2 <- lme(log(Pb)~Age+Depth,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(Pbmodsim,Pbmodsim2)
Pbmodsim3 <- lme(log(Pb)~Age,random=~1|Cell_no,data=fullsoil,na.action=na.omit,method="ML")
anova(Pbmodsim2,Pbmodsim3)
summary(Pbmodsim2)
  #The interaction between depth and age is really close (p=0.07) so may want to maintain

#Test for significance of the random effect
Pbmod_noME <- lm(log(Pb)~Age*Depth,data=fullsoil)
anova(Pbmodsim,Pbmod_noME)
  #Yes, retain random effect
```

Pb concentrations increase with age, and there is marginal trend of greater increases in the surface soil when compared to deep soils (Age * Depth, p=0.07). Seems like Pb is an intermediate between Cu and Zn where Zn has no vertical stratification, Cu is strongly stratified and Pb is weakly stratified. Pb increases 4-12% (deep and surface, respectively) year-over-year.

##Other variables influencing Cu

```{r}

#Test models with new variables

#STEP 1, including one additional
candvar <- c('Cell_area','OM','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logOM','logBD','logd50','logDrain','logCell','logDtoC')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Pbresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Pbdata <- fullsoil[,c(1,4,6,17,loc[i])]
  Pbdata <- Pbdata[complete.cases(Pbdata),]
  
  MPb1 <- lmer(log(Pb)~Age+Depth+(1|Cell_no),data=Pbdata,REML=F)
  MPb2 <- lmer(log(Pb)~Age+Depth+Pbdata[[5]]+(1|Cell_no),data=Pbdata,REML=F)
  
  Pbresult$Pval[i] <- anova(MPb2,MPb1)$P[2]
}

Pbresult
#INCLUDE logOM (p<0.001).

#REFINE THE Pb MODEL
MPb2 <- lmer(log(Pb)~Age+Depth+logOM+(1|Cell_no),data=fullsoil,REML=F)
MPb2.1 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Depth+(1|Cell_no),data=fullsoil,REML=F)
anova(MPb2,MPb2.1)

MPb2.2 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
anova(MPb2.2,MPb2)
  #Significant age x log(OM) interaction

#STEP 2 - See if other variables besides OM should be included
candvar <- c('Cell_area','BD','d50','Drain_area','DtoC','Kfs','LF_sus','HF_sus',
             'logCell','logDrain','logDtoC','logBD','logd50')
loc <- which(names(fullsoil) %in% candvar)
names(fullsoil)[loc]

Pbresult <- data.frame("Variable"=names(fullsoil)[loc],"Pval"=NA)

for(i in 1:length(candvar)){
  Pbdata <- fullsoil[,c(1,4,6,14,17,loc[i])]
  Pbdata <- Pbdata[complete.cases(Pbdata),]
  
  MPb1 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=Pbdata,REML=F)
  MPb2 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+Pbdata[[6]]+(1|Cell_no),data=Pbdata,REML=F)
  
  Pbresult$Pval[i] <- anova(MPb2,MPb1)$P[2]
}

Pbresult
#No other variables are worth including.

#FINAL MODEL
MPb2.2 <- lmer(log(Pb)~Age+Depth+logOM+logOM:Age+(1|Cell_no),data=fullsoil,REML=F)
summary(MPb2.2)
plot(MPb2.2)

MPb2nlme <- lme(log(Pb)~Age+Depth+logOM+Age:logOM,random=~1|Cell_no,data=fullsoil,na.action=na.omit)
anova(MPb2nlme)

#Predictions to try and interpret the model
newage <- seq(0,7)
highOM <- predict(MPb2.2,newdata=list(Age=newage,Depth=rep("0 to 10",8),logOM=rep(log10(80),8)),re.form=~0)
lowOM <- predict(MPb2.2,newdata=list(Age=newage,Depth=rep("0 to 10",8),logOM=rep(log10(5),8)),re.form=~0)

plot(Pb~Age,data=fullsoil)
lines(newage,exp(highOM))
lines(newage,exp(lowOM))
```

Pb in bioretention soils increase through time, but only in soils that had higher OM content (Age x log(OM) p = 0.0001). Low OM content soils showed no accumulation of Pb through time, but high OM soils accumuated substantial Pb (20-150 mg/kg over a 7 year period). Deep soils overall had less Pb (p=0.016) but even deep soils with high OM accumulated Pb over time.
